---
title: "Quiz 12 Solution"
author: "Lucas Carlos"
date: "November 13, 2017"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE)
knitr::opts_chunk$set(comment = '')
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error   = FALSE)
```

# Introduction

This quiz introduces you to some basic exploratory data analysis using tools from the tidyverse collection of packages.  I have provided the solution as a pdf.  Your job is to fill in the missing code where you see blanks and experiment with Rmarkdown to learn about some of the functions.  Good luck!  I look forward to seeing your finished files on our canvas page

# Load packages using library()

```{r}
library(tidyverse)
library(tidyr)
```


# Use head() to view the first 5 lines of mtcars

```{r}
head(mtcars, n = 5)
```

# Use names() to find the columns (variable) name)

```{r}
names(mtcars)
```

# Use dim() to find the number of rows and columns of mtcars

```{r}
dim(mtcars)
```

# Use glimpse() to take a look at mtcars

```{r}
glimpse(mtcars)
```

# Use str() to view the structure of mtcars

```{r}
str(mtcars)
```

# Use summary() to make numeric summary of mtcars

```{r}
summary(mtcars)
```

# Use the pipe operator, group_by(), and summarize() to find the average mpg for each different value of cyl

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(avg_mpg = mean(mpg))
```

# For each different value of gear, find the average mpg

```{r}
mtcars %>% 
  group_by(gear) %>% 
  summarize(avg_mpg = mean(mpg))
```

# Plot mpg vs hp and mapping gear to color ggplot(), geom_point(), ggtitle

```{r}
mtcars %>% 
  ggplot(aes(x = hp, y = mpg, col = as.factor(gear))) +
  geom_point() +
  ggtitle("mpg vs. hp")
```

# Create three side by side plot for each value of gear using facet_wrap()

```{r}
mtcars %>% 
  ggplot(aes(x = hp, y = mpg)) +
  geom_point(aes(col = as.factor(gear))) +
  ggtitle("mpg vs. hp") +
  facet_wrap(~gear)
```

