---
title: "Quiz 13 Solution"
author: "Lucas Carlos"
date: "November 13, 2017"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE)
knitr::opts_chunk$set(comment = '')
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error   = FALSE)
```

## Introduction

This quiz introduces you to some basic exploratory data analysis using tools from the tidyverse collection of packages.  I have provided the solution as a pdf.  Your job is to fill in the missing code where you see blanks and experiment with Rmarkdown to learn about some of the functions.  Good luck!  I look forward to seeing your finished files on our canvas page

## Loading and Installing Packages

Load packages using library().  You may need to install these packages first if they are not already installed

```{r}
library(tidyverse)
library(tidyr)
library(gapminder)
```


## Getting to know a dataset

Use head() to view the first 5 lines of gapminder

```{r}
head(gapminder, n = 5)
```

Use names() to find the columns (variable) name)

```{r}
names(gapminder)
```

Use dim() to find the number of rows and columns of gapminder

```{r}
dim(gapminder)
```

Use glimpse() to take a look at gapminder

```{r}
glimpse(gapminder)
```

Use str() to view the structure of gapminder

```{r}
str(gapminder)
```

Use summary() to make numeric summary of gapminder

```{r}
summary(gapminder)
```

## Selecting a subset of the Variables

You use select() when you want to select a subset of the variables.  Use select() to select only the continent, year, and pop.

```{r}
gapminder %>%
  select(continent, year, pop)
```


## Filtering Observations 

Add a filter() line after the pipe (%>%) to extract only the observations from the year 1957.   Remember that you use == to compare two values.  

```{r}
gapminder %>% 
  filter(year == 1957) 
```



Filter the gapminder data to retrieve only the observation from China in the year 2002.

```{r}
gapminder %>% 
  filter(year == 2002, country == "China") 
```

# Arranging Observations

You use arrange() to sort observations in ascending or descending order of a particular variable. In this case, you'll sort the dataset based on the pop variable.

```{r}
# Sort in ascending order of lifeExp
gapminder %>%
  arrange(pop)
  
# Sort in descending order of lifeExp
gapminder %>%
  arrange(desc(lifeExp))
```

## Filtering and arranging

You will use the pipe operator (%>%) to combine multiple dplyr verbs in a row. In this case, you'll combine a filter() with an arrange() to find the highest population countries in a particular year. Use filter() to extract observations from just the year 1962, then use arrange() to sort in descending order of population (pop).

```{r}
gapminder %>% 
  filter(year == 1962) %>%
  arrange(desc(pop))
```

## Using mutate to create new column

Suppose we want life expectancy to be measured in months instead of years: you'd have to multiply the existing value by 12. You can use the mutate() verb to change this column, or to create a new column that's calculated this way.

```{r}
# Use mutate to change lifeExp to be in months
gapminder %>%
  mutate(lifeExp = lifeExp * 12)

# Use mutate to create a new column called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = lifeExp * 12)
```

## Combining filter, mutate, and arrange

In this exercise, you'll combine all three of the verbs you've learned in this chapter, to find the countries with the highest life expectancy, in months, in the year 2007.

```{r}
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = lifeExp * 12) %>%
  arrange(desc(lifeExpMonths))
```

## Using group_by and summarize()

You can group data using the group_by() function and create summary information with summarize.  Use filter() to select only observations from the year 1972, then use group_by() to group the by continent, then use summarize() to calaculate the mean life expectancy by continent and use arrange() to arrange continents by descending order of life expectncy.

```{r}
gapminder %>%
  filter(year == 1972) %>%
  group_by(continent)  %>%
  summarize(mean_exp = mean(lifeExp)) %>%
  arrange(desc(mean_exp))
```

