---
title: "Quiz 15 Introduction to Matrices in R"
author: "Lucas Carlos"
date: "November 28, 2017"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE)
knitr::opts_chunk$set(comment = '')
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error   = FALSE)
library(tidyverse)
```

## Introduction

Matrix Algebra is very important in applied mathematics.  This Quiz walks you through some basic matrix fucntions in R.  To get started lets create two 4 x 4 matrices called A and B respectively


## Create a matrix by rows

Use rbind() to create a 4 x 4 matrix

```{r}
library(dplyr)
library(tibble)
A <- rbind(
  c(1, 2, 6, 4),
  c(5, 1, 7, 8),
  c(9, 8, 7, 6),
  c(2, 4, 3, 0))

A
```

The above method of creating a matrix has the advantage that what you see is what you get.  

## Create a matrix by binding columns, cbind()

```{r}
B <- cbind(
  c(6, 7, 1, 0),
  c(0, 3, 5, 7),
  c(2, 4, 6, 4),
  c(9, 1, 3, 6))

B
```

)

## The transpose of a matrix

Find the transpose of A using ()

```{r}
A

t(A)
```
## Row means, colmeans, col sums and row sums

Can quickly apply function by columns or rows using apply().  Use MARGIN = 1 for row and MARGIN = 2 for columns.  Also try rowSums(), colSums(), rowMeans(), and colMeans()

```{r}
# Row sums
apply(X = A, MARGIN = 1, FUN = sum)
rowSums(A)

# Column Sums
apply(X = A, MARGIN = 2, FUN = sum)
colSums(A)

# Row means
apply(X = A, MARGIN  = 1, FUN = mean)
rowMeans(A)

# Column Means
apply(X = A, MARGIN = 2, FUN = mean)
colMeans(A)


# Row standard deviation
apply(X = A, MARGIN = 1, FUN = sd)


# Column standard deviation
apply(X = A, MARGIN = 2, FUN = sd)

```


## Adding two matrices in R

Add the matrices together, assign to C using the assignment operator

```{r}
C <- A + B
C
```

## Subtracting two matrices

Subtract the matrices from each other, assign to D using the assignment operator <-

```{r}
D <- A - B
D
```

## Multiplying componentwise

If you use * between two matrice of the same dimension you get elementwise products. 

```{r}
A * B
```

## Matrix multiplication

Matrix multiplication is not equal to element wise multiplication.  Matrix multiplication is achieved with %*%

```{r}
A %*% B
# Notice that matrix multiplication is not commutative
B %*% A
# Matrix multiplication is associative 
(A %*% B) %*% D
A %*% (B %*% D)
```

## Subsetting Matrices

You can select specific matrix entries as follows

```{r}
# Select the entry of A in the second row and third column
A[2, 3]

# Select the third row of B
A[3, ]

# Select the 1st column
A[, 1]

# Select the diagonal of A using diag()
diag(A)
```

## Adding row and column names

You can add row names and column names to a matrix using colnames() and rownames()

```{r}
colnames(B) <- c('blue', 'orange', 'gold', 'purple')
B

rownames(B) <- c('hot', 'cold', 'spicy', 'sweet')
B
```

## Subsetting a matrix using names

You can use row and column names to select specific entries

```{r}
# Select the hot row of B
B["hot", ]

# Select the gold columns of B
B[, "gold"]

# Select the spicy and purple entry
B["spicy", "purple"]
```

## Converting a matrix to a data frame and shifting rownames 

```{r}
B_frame <- B %>% 
  as.data.frame() %>%
  rownames_to_column(var = "flavor")

B_frame
```

## Using ggplot to visualize data

Once you have converted a matrix to a data frame you can use ggplot to graph data.  Use cex = 4 inside of geom_point() to make the dots larger.

```{r}
B_frame %>%
  ggplot(aes(x = gold, y = blue, color = flavor)) +
  geom_point(cex = 4)
```

